---
title: "STAT 206 Homework 1"
subtitle: 'ChiKit Ho'
output: pdf_document
---

**Due Monday, October 12, 5:00 PM**

***General instructions for homework***: Homework must be submitted as pdf file, and be sure to include your name in the file.  Give the commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file. Each answer must be supported by written statements as well as any code used.  (Examining your various objects in the "Environment" section of RStudio is insufficient -- you must use scripted commands.)

1. The data set at [http://www.faculty.ucr.edu/~jflegal/206/rnf6080.dat] records hourly rainfall at a certain location in Canada, every day from 1960 to 1980.
    a. First, we need to load the data set into R using the command `read.table()`. Use the help function to learn what arguments this function takes. Once you have the necessary input, load the data set into R and make it a data frame called `rain.df`.
```{r}
rain.df <- read.table('http://www.faculty.ucr.edu/~jflegal/206/rnf6080.dat')
```
    b. How many rows and columns does `rain.df` have? (If there are not 5070 rows and 27 columns, something is wrong; check the previous part to see what might have gone wrong in the previous part.)
```{r}
dim(rain.df)
```
    c. What are the names of the columns of `rain.df`?
```{r}
names(rain.df)
```
    d. What is the value of row 5, column 7 of `rain.df`?
```{r}
rain.df[5,7]
```
    e. Display the second row of `rain.df` in its entirety.
```{r}
rain.df[2,]
```
    f. Explain what this command does:
```
names(rain.df) <- c("year","month","day",seq(0,23))
```
by running it on your data and examining the object. (You may find the display functions `head()` and `tail()` useful here.) Is it clear now what the last 24 columns represent?
```{r}
names(rain.df) <- c("year","month","day",seq(0,23))
head(rain.df)
tail(rain.df)
```
It names the fist column as year,the second column as month and the third column as day
The rest of the 24 columns represents the 24 hours for the date accordingly.
   
    g. Create a new column in the data frame called `daily`, which is the sum of the rightmost 24 columns. With this column, create a histogram of the values in this column, which are supposed to be daily rainfall values. What is wrong with this picture? 
    
```{r}
rain.df$daily <-rowSums(rain.df[,4:27])
hist(rain.df$daily)
```
There are some columns with the negative number of daily rainfall, which is not possible

    h. Create a new data frame `rain.df.fixed` that takes the original and fixes it for the apparent flaw you have discovered. Having done this, produce a new histogram with the corrected data and explain why this is more reasonable.
    
```{r}
rain.df.fixed<-rain.df
rain.df.fixed[rain.df.fixed<0]<-NA
rain.df.fixed$daily <- rowSums(rain.df.fixed[,4:27], na.rm=TRUE)
hist(rain.df.fixed$daily)
```

2. Syntax and class-typing.
    a. For each of the following commands, either explain why they should be errors, or explain the non-erroneous result.

```{r, eval=FALSE}
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
sum(vector1)
```
**Answer**
The elements in vector1 are entered as characters instead of number. So, the result of Max(vector1) & sort(Vector1) are funny and there is an error message for sum(vector1). To solve the errors we can enter the vector1<-c(5,12,7,32) without the quoatation so the elements of vector1 are intergers.

  b. For the next series of commands, either explain their results, or why they should produce errors.

```{r, eval=FALSE}
vector2 <- c("5",7,12)
vector2[2] + vector2[3]

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]
list4[2]+list4[4]
```

**Answer**
The elements in vector2 are treated as characters, so, vector2[2]+vector2[3] will result in error becuase they are characters. To solve it, we can reenter vector2 as vector2<-c(5,7,12).
The dataframe treated the elements as integer, so, dataframe3[1,2] + dataframe3[1,3] will result as 19
list4 is a list with 2 charaters and 2 integers. list4[[2]]+list4[[4]] will result in 168 while list4[2]+list4[4] will result in errors
because there is only 1 [].

3. Working with functions and operators.
    a. The colon operator will create a sequence of integers in order. It is a special case of the function `seq()` which you saw earlier in this assignment. Using the help command `help(seq)` to learn about the function, design an expression that will give you the sequence of numbers from 1 to 10000 in increments of 372. Design another that will give you a sequence between 1 and 10000 that is exactly 50 numbers in length.
```{r}
seq(1,10000,by=372)
seq(1,10000,length.out=50)
```
    b. The function `rep()` repeats a vector some number of times. Explain the difference between `rep(1:3, times=3) and rep(1:3, each=3).
```{r}
rep(1:3, times=3)
rep(1:3, each=3)
```
rep(1:3, times=3) repeat 1 to 3 in 3 times while 
rep(1:3, each=3) repeat each of numbers 3 times.
